Velocity Map Imaging Spectroscopy of C$_2$H$^-$ and C$_2$D$^-$: a benchmark study of vibronic coupling interactions

### 2022-06-24 Resubmission/Laws_C2H.tex
update resubmission

\documentclass[journal=jpcafh,manuscript=article,layout=onecolumn, 12pt]{achemso}
\usepackage[version=3]{mhchem} % Formula subscripts using \ce{}
\usepackage[T1]{fontenc}       % Use modern font encodings
\usepackage{amsmath}
\usepackage[dvipsnames]{xcolor}
% NB added command for in line cite
\newcommand{\onlinecite}[1]{\hspace{-1 ex} \nocite{#1}\citenum{#1}} 
%
\author{Benjamin~A.~Laws}
\email{b.laws@unsw.edu.au}
\affiliation{School of Chemistry, University of New South Wales, Sydney NSW 2052, Australia}
\alsoaffiliation{Research School of Physics, The Australian
	National University, Canberra ACT 2601, Australia}
\author{Zachariah~D.~Levey} 
\affiliation{School of Chemistry, University of New South Wales, Sydney NSW 2052, Australia}
\author{Andrei Sanov}
\affiliation{Department of Chemistry and Biochemistry, The University of Arizona, Tucson, Arizona 85721, United States}
\author{John F. Stanton}
\affiliation{Department of Chemistry, University of Florida, Gainesville, Florida 32611, United States}
\author{Timothy~W.~Schmidt} 
\affiliation{School of Chemistry, University of New South Wales, Sydney NSW 2052, Australia}
\author{Stephen~T.~Gibson}
\affiliation{Research School of Physics, The Australian
	National University, Canberra ACT 2601, Australia}
\title{Velocity Map Imaging Spectroscopy of C$_2$H$^-$ and C$_2$D$^-$: a benchmark study of vibronic coupling interactions}
\abbreviations{PES,PAD,EA,eKE,FWHM,VMI}
\begin{document} 
\begin{abstract}

\end{abstract}
High-resolution velocity map imaged photoelectron spectra of the ethynyl anions C$_2$H$^-$ and C$_2$D$^-$ are measured over a range of photodetachment wavelengths, to investigate the complex interactions between the close lying $\tilde{X}^2\Sigma^+$ and $\tilde{A}^2\Pi$ electronic states. Vibronic coupling calculations are performed by transforming to a quasidiabatic representation, parametrised by \emph{ab-initio} CFOUR calculations, to model the interplay between pseudo Jahn-Teller and Renner-Teller coupling. This approach is combined with electron anisotropy measurements to assign all of the observed vibronic structure, providing a framework that may be used to accurately simulate spectra of larger C$_{2n}$H monohydride carbon chains.

@@

Chain growth of the C$_{2n}$H radicals occurs predominately through acetylene addition,~\cite{wil03}
\begin{equation}
\text{C}_{2n}\text{H} + \text{C}_2\text{H}_2 \rightarrow \text{C}_{2n+2}\text{H}+\text{H}_2.
\end{equation}
Conversely, the formation of anions in space is believed to be driven by radiative electron attachment, charge transfer, and dissociative electron attachment. Current astrochemical modelling of interstellar clouds, such as IRC+10216, typically underestimate the observed abundances of CH$_4^-$~\cite{mil17,cor13,her08}. Therefore, it has been proposed that ion-neutral reactions
 \begin{equation}
	\text{C}_{2n}\text{H}^- + \text{C}_2\text{H}_2 \rightarrow \text{C}_{2n+2}\text{H}^- + \text{H}_2.
\end{equation}
may be significant chain growth mechanisms that also needs to be included in astrochemical models~\cite{mil17,bas19}. Studies of the extraterrestrial planetary atmosphere of Saturn's moon Titan suggest that both C$_2$H and C$_2$H$^-$ play an important role in the atmospheric chemistry~\cite{dob16,vui09,des17,vri18}. Multiple reaction pathways involving the C$_2$H$^-$ have been proposed, including reactions with C$_2$H$_2$ and HCN.

@@

In this work the vibronic spectrum of the ethynyl radical C$_2$H is examined with High-Resolution Photoelectron Imaging (HR-PEI) together with CFOUR~\cite{dev20}  \emph{ab-initio} vibronic-coupling calculations. C$_2$H is reported to be one of the most abundant molecules in the universe, and is the most thoroughly studied of the C$_{2n}$H species~\cite{wil91,hei99,for10}. Many different experimental techniques have been employed to help understand the complexities of the vibronic spectrum, including electron spin resonance~\cite{coc64,jin85}, laser magnetic resonance~\cite{bro88,pfe96,sch98}, microwave and millimeter wave spectroscopy~\cite{sas81,got83,end89,mul00}, infrared (matrix isolation~\cite{she87,jac87,for95} and Fourier Transform emission~\cite{ver88}) spectroscopy, laser induced fluorescence spectroscopy~\cite{hsu93,hsu95,chi99} and photoelectron spectroscopy of the negative ion~\cite{erv91,tay98,zho07}. C$_2$H has also received extensive theoretical attention to interpret the experimental results, however the large number of possible vibronic levels means this has proven to be a challenge~\cite{per90,per91,per91b,per91c,per92,car00,tar03,tar04,for10,sta21,gul21}. In this work we demonstrate how the construction of a quasidiabatic Hamiltonian allows the strength of vibronic interactions between coupled surfaces near a conical intersection to be estimated, sufficient to accurately simulate electronic and vibronic spectra. %The quadratic vibronic coupling model employed can be readily scaled to larger molecules (such as longer C$_n$H chains) which quickly become too complex to study using full variational approaches. 
Anion HR-PEI spectroscopy maps both of the $^2\Sigma^+$ and $^2\Pi$ surfaces on an equal footing from the anion $\tilde{X}^1\Sigma^+$ state, allowing a direct comparison with \emph{ab-initio} modelling. %The electron anisotropy allows transitions with $\sigma$ and $\pi$ symmetry to be distinguished, providing definitive assignments of calculated transitions.

@@

In C$_2$H the loss of degeneracy creates a Renner-Teller (RT) pair in the excited state, where the usually degenerate $\Pi$ surfaces separate to form two non-degenerate electronic states $\Pi^+ (2A')$ and $\Pi^-(1A'')$. This involves separating a single potential energy surface (V) into two distinct but coupled surfaces (V$^+$) and (V$^-$). Due to the strong coupling along the linear axis between the electronic and vibration angular momenta of the $2A'$ and $1A''$ components of the $^2\Pi$ state, stationary states cannot be explicitly assigned to either of the $\Pi^+(2A')$ or $\Pi^-(1A'')$ electronic surfaces. Instead, they exist as a combination of both states. Due to the close lying nature of the ground {$\tilde{X} ^2\Sigma^+$} and excited {$\tilde{A} ^2\Pi$} electronic states, which are only separated by $\sim3700~$cm$^{-1}$, a pseudo Jahn-Teller effect is also observed. %The Jahn-Teller theorem states that stability and degeneracy are not possible simultaneously unless the molecule is linear~\cite{jah37}, meaning that non-linear molecules with degenerate electronic states will undergo a symmetry breaking distortion in order to remove the degeneracy. However a similar effect has also been observed where coupling exists between a non-degenerate state and a nearby pair of degenerate states, even if a molecule is linear. 
In the case of C$_2$H this is seen as coupling between the ground $\Sigma^+(1A')$ and excited $\Pi^+(2A')$ states, induced by the bending motion of $v_2$. The ground state only couples to one member of the Renner-Teller pair $\Pi^+(2A')$, as the other state $\Pi^-(1A'')$ has incompatible symmetry. This results in the complex vibronic structure observed for the $\tilde{A}^2\Pi$ electronic state, with contributions from three coupled surfaces $\Sigma^+(1A'),$ $\Pi^+(2A'),$ and $\Pi^-(1A'')$.

@@

\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth]{figures/Fig3}
	\caption{Adiabatic potential energy surfaces of $\tilde{X}^2\Sigma^+$, $\tilde{A}^2\Pi^-$, and $\tilde{A}^2\Pi^+$ states, calculated using the parameters calculated by Tarroni and Carter~\cite{tar03}.}
	\label{fig:2}
\end{figure}

The adiabatic potential energy surfaces in Figure~\ref{fig:2} were evaluated from the variational parameters of Tarroni and Carter~\cite{tar03}, illustrating the loss of degeneracy for the doublet $\Pi$ surfaces, due to the Renner-Teller and pseudo Jahn-Teller interactions. This technique has been used to calculate a multitude of admixed vibronic levels. More than 100 possible C$_2$H vibronic states have been calculated, up to 10,000~cm$^{-1}$ above the $\tilde{X} ^2\Sigma^+$ origin. However, assigning experimental spectra has remained a challenge, partly due to the large number of calculated levels providing for multiple possible assignments for each experimentally observed transition.

\subsection{Coupling Calculations}
In order to guide spectral assignment of C$_2$H and future astronomical searches for other C$_n$H radicals, transition intensities are also needed. However, due to the strong vibronic coupling interactions, transition intensities can not be obtained via simple Franck-Condon calculations based on quantum chemistry methods. To account for these interactions, the photoelectron spectrum of the C$_2$H$^-$ anion is simulated using a quasidiabatic Hamiltonian of the type advocated by Koeppel, Domcke and Cederbaum (KDC).~\cite{kou84,dom81} In this approach, the Hamiltonian is represented in a basis of quasidiabatic (slowly varying) electronic  states for which the kinetic energy operator can be assumed diagonal.~\cite{pac93} For C$_2$H, the KDC Hamiltonian comprises three states - the two components of the $^2\Pi$ state and the ground $^2\Sigma^+$ state - and is then projected onto a vibrational basis, usually chosen as a direct product of harmonic oscillators. Diagonalization of the corresponding matrix yields the molecular states (which are given in terms of a Born-Huang expansion), and the squared projections of the corresponding eigenvectors onto the ground state of the anion yield the relative intensities. The latter is true only if the photodetachment cross sections for the $^2\Sigma^+$ and $^2\Pi$ states are assumed equal, but different cross sections of the two states can be incorporated by scaling the intensities of states according to their (vibronic) symmetry.~\cite{sta11}

@@

When the potential in Eq.~(\ref{eq:calc1}) is diagonalized, the adiabatic states that are used for its parametrization are precisely recovered through terms second order in displacement. The linear diabatic coupling constants between the electronic states of interest were calculated analytically at the EOM-CCSD/ANO2 level of theory, using the CFOUR computational package~\cite{dev20} and the quasidiabatic ansatz of Ichino~\emph{et al.}~\cite{ich09} Briefly, the EOM procedure is used to operationally define quasidiabatic states (those that relax according to a well-behaved reference state wavefunction, which is that of the anion level) and the coupling constants are then evaluated as the first derivative of the off-diagonal elements of the electronic Hamiltonian in the basis defined by this representation. The diabatic force and coupling constants calculated to parametrize the KDC potential in Eq.~(\ref{eq:calc1}) were calculated at the CCSDT/ANO2 level of theory, and are presented in Table~\ref{tab:1}.

%To parametrize this Hamiltonian, the equation-of-motion coupled-cluster method known as EOMIP-CCSDT has been used to optimise the geometry of the ground state anion, with an Atomic Natural Orbital (ANO) basis (ANO2), using the CFOUR computational package.

%Adiabatic linear and quadratic force constants are calculated from the analytical derivatives of the \emph{ab-initio} potential energy surface with respect to the anion reduced normal coordinates, at the anion geometry Q$_0$. The interstate PJT coupling constant $\lambda$ is evaluated directly using the CFOUR package to employ the quasidiabatic ansatz described in refs. This allows for direct ab-initio evaluation of the linear diabatic interstate coupling constants that parametrize the KDC potential. The calculated parameters are presented in the Supplementary Materials, in Table S1.

\begin{table}
	\caption{Parameters for the quasidiabatic Hamiltonian Eq.~(\ref{eq:calc1}) determined using CFOUR and the quasidiabatic ansatz. Units are in cm$^{-1}$.}
	\label{tab:1}
%	\begin{tabular}{c |c c c c| c c c c }
%		 & \multicolumn{4}{c}{C$_2$H}     & \multicolumn{4}{c}{C$_2$D}  \\ \hline
%		  & Anion & $\tilde{X}^2\Sigma^+$ & $\tilde{A}^2\Pi (a')$ & $\tilde{A}^2\Pi (a'')$ & Anion & $\tilde{X}^2\Sigma^+$ & $\tilde{A}^2\Pi (a')$ & $\tilde{A}^2\Pi (a'')$ \\
%		 \hline\hline
%		 $\Delta$E & 0 & 22431.21 &  26543.67 & 26543.67 & 0&	22431.21 & 26543.67	& 26543.67\\
%		 F$_1$ & 0	& -211.28 & 286.9 &	286.9 & 0& -596.04& 641.6 & 641.6\\
%		 F$_3$ & 0 & -1515.21 & 1257.07 & 1257.07 & 0 & -1401.53 & 1115.17 & 1115.17\\
%		 F$_{2a}$ & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
%		 F$_{2b}$ & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
 %        F$_{11}$ & 3424.32 & 3487.03 & 3482.16 & 3482.16 & 2611.42 & 2636.57 & 2681.25 & 2681.25\\
%         F$_{13}$ & 0.00 & -28.32 & 31.46 & 31.46 & 0.00 & -48.38 & 55.40 & 55.40\\
 %        F$_{33}$ & 1894.71 & 1830.83 & 2099.57 & 2042.21 & 1792.49 & 1747.62 & 1914.39 & 1914.39\\
  %       F$_{2a2a}$ & 607.09 & 87.57 & 1883.48 & 658.96 & 481.89 & 120.09 & 1503.88 & 473.55\\
   %      F$_{2b2b}$ & 607.09 & 87.57 & 658.96 & 1883.48 & 481.89 & 120.09 & 473.55 & 1503.88\\
    %     $\lambda$ &  & & -1185.26 &  &  & & -1014.58 &  \\
     %    \end{tabular}
     \begin{tabular}{c | c c c c | c c c c} 
     	\hline
     	& \multicolumn{4}{c}{C2H} & \multicolumn{4}{c}{C2D}\\
     	\hline
     	& Anion & $\tilde{X}^2\Sigma^+$ & $\tilde{A}^2\Pi (a')$ & $\tilde{A}^2\Pi (a'')$ & Anion & $\tilde{X}^2\Sigma^+$ & $\tilde{A}^2\Pi (a')$ & $\tilde{A}^2\Pi (a'')$  \\
     	\hline
     	$\Delta_0$ & 0& 22431.21&  26543.67& 26543.67& 0 & 22431.21& 26543.67& 26543.67 \\
     	F$_1$ &  0& -211.28& 286.90& 286.90& 0& -596.04& 641.6& 641.6\\
     	F$_3$ &  0& -1515.21&  1257.07& 1257.07& 0& -1401.53& 1115.17& 1115.17\\
     	F$_{2a}$& 0 & 0& 0& 0& 0& 0& 0&0 \\
     	F$_{2b}$&  0& 0& 0& 0& 0& 0& 0&0 \\
     	F$_{11}$& 3424.32 &  3487.03& 3482.16& 3482.16& 2611.42&  2636.57& 2681.25& 2681.25\\
     	F$_{13}$&  0 & -28.32& 31.46& 31.46& 0& -48.38& 55.40& 55.40\\
     	F$_{33}$& 1894.71 & 1830.83& 2042.21& 2042.21& 1792.49&  1747.62& 1914.39& 1914.39\\
     	f$_{22}$& 607.09&  87.57& -& -& 481.89&  120.09& -& -\\
     	F$_{22}$& 607.09 & 770.78& -& -& 481.89& 620.70& -& -\\
     	f$_{2a2a}$& - &- & 1883.48& 658.96& -& -& 1503.88& 473.55\\
     	f$_{2b2b}$& - &- & 658.96& 1883.48& -& -& 473.55& 1503.88\\
     	F$_{2a2a}$& -& -& 1200.27& 658.96& -& -& 1003.27& 473.55\\
     	F$_{2b2b}$& -& -& 658.96& 1200.27& -& -& 473.55& 1003.27\\
     	&&&&&&&& \\
     	$\lambda$& & & -1185.26& & & & -1014.58& \\
     	$\eta$ & & & -270.66& & & & -264.86& \\

@@

\subsection{Simulated Spectra}
The above method was employed to simulate the photoelectron spectrum of C$_2$H$^-$, with the calculated transitions shown in Figure~\ref{fig:3}(a) alongside the experimental data at 355~nm. Photoelectron spectra are simulated using the xsim package of CFOUR and the model Hamiltonian generated above. The Hamiltonian is diagonalised using the Lanczos algorithm, to calculate transition energies and intensities that map to the measured photoelectron spectrum.  %Calculated transitions with $\Sigma$ symmetry are in blue, and transitions with $\Pi$ symmetry are in orange. 
On the $\tilde{X} ^2\Sigma^+$ surface, below 3,600~cm$^{-1}$, there is excellent agreement in both the transition positions and intensities between the simulated and experimental spectra. This includes the HT coupled (2$^{n+1}$) transitions with $\Pi$ symmetry, which would normally be missing from a simulation using the standard \emph{ab-initio} approaches.

@@

\begin{figure}[th!]
	\includegraphics[width=0.6\textwidth]{figures/Fig5.pdf}
	\caption{(a) Experimental photoelectron spectrum of C$_2$H$^-$ at 355~nm, compared with the simulated spectrum, calculated using the quasidiabatic ansatz. Transitions with $\Sigma$ symmetry are shown in blue, and those with $\Pi$ symmetry are shown in orange. (b) Experimental photoelectron spectrum of C$_2$D$^-$ at 355~nm, compared to the simulated spectrum. The labelling of peaks near the $\tilde{A}$ origin correspond to transitions identified in Tables~S1 and~S2.}
	\label{fig:3}
\end{figure}

@@

The simulated spectrum reproduces the large change observed for the deuterated species. The calculated intensity pattern of the vibronic levels around the $\tilde{A} ^2\Pi$ origin also mirrors the experimental data, with a single intense transition observed near peak $e$, and 3 prominent transitions predicted around the experimental peaks $a-d$. The calculated splitting between the levels is also similar. %however the relative intensity of peak $d$ appears to be underestimated. %The results from Figure~\ref{fig:C2H-plot} confirm that the quasiadiabtic approach is able to accurately describe the vibronic interactions between the $\Sigma$ and $\Pi$ surfaces, including deuteration effects.

@@

%Overall, figures a-c show that the quasidiabatic approach is a reliable method for simulating the spectra of molecules when strong vibornic coupling interactions are involved. This method may be expanded to other C$_n$H radicals that possess similar complexities, in order to help guide the search for new molecules to be found in space.

>>>>

%% ****** Start of file aiptemplate.tex ****** %
%%
%%   This file is part of the files in the distribution of AIP substyles for REVTeX4.
%%   Version 4.1 of 9 October 2009.
%%
%
% This is a template for producing documents for use with 
% the REVTEX 4.1 document class and the AIP substyles.
%

\documentclass[aip,graphicx]{revtex4-1}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage[version=3]{mhchem} % Formula subscripts using \ce{}
\usepackage[T1]{fontenc}       % Use modern font encodings
\usepackage{amsmath}
\usepackage[dvipsnames]{xcolor}
% NB added command for in line cite
%\newcommand{\onlinecite}[1]{\hspace{-1 ex} \nocite{#1}\citenum{#1}} 
%\documentclass[aip,reprint]{revtex4-1}

\draft % marks overfull lines with a black rule on the right

\begin{document}

% Use the \preprint command to place your local institutional report number 
% on the title page in preprint mode.
% Multiple \preprint commands are allowed.
%\preprint{}

\title{Velocity Map Imaging Spectroscopy of C$_2$H$^-$ and C$_2$D$^-$: a benchmark study of vibronic coupling interactions}

\author{Benjamin~A.~Laws}
\email{b.laws@unsw.edu.au}
\affiliation{School of Chemistry, University of New South Wales, Sydney NSW 2052, Australia}
\affiliation{Research School of Physics, The Australian
	National University, Canberra ACT 2601, Australia}
\author{Zachariah~D.~Levey} 
\affiliation{School of Chemistry, University of New South Wales, Sydney NSW 2052, Australia}
\author{Andrei Sanov}
\affiliation{Department of Chemistry and Biochemistry, The University of Arizona, Tucson, Arizona 85721, United States}
\author{John F. Stanton}
\affiliation{Department of Chemistry, University of Florida, Gainesville, Florida 32611, United States}
\author{Timothy~W.~Schmidt} 
\affiliation{School of Chemistry, University of New South Wales, Sydney NSW 2052, Australia}
\author{Stephen~T.~Gibson}
\affiliation{Research School of Physics, The Australian
	National University, Canberra ACT 2601, Australia}

\date{\today}

\begin{abstract}
High-resolution velocity-map imaged photoelectron spectra of the ethynyl anions C$_2$H$^-$ and C$_2$D$^-$ are measured over a range of photodetachment wavelengths, to investigate the complex interactions between the close lying $\tilde{X}^2\Sigma^+$ and $\tilde{A}^2\Pi$ electronic states. Vibronic coupling calculations are performed by transforming to a quasidiabatic representation, parametrised by \emph{ab-initio} CFOUR calculations, to model the interplay between pseudo Jahn-Teller and Renner-Teller coupling. This approach is combined with electron anisotropy measurements to assign all of the observed vibronic structure, providing a framework that may be used to accurately simulate spectra of larger C$_{2n}$H monohydride carbon chains.
\end{abstract}

\pacs{}

\maketitle

@@

Chain growth of the C$_{2n}$H radicals occurs predominately through acetylene addition,~\cite{wil03}
\begin{equation}
\text{C}_{2n}\text{H} + \text{C}_2\text{H}_2 \rightarrow \text{C}_{2n+2}\text{H}+\text{H}_2.
\end{equation}
Conversely, the formation of anions in space is believed to be driven by radiative electron attachment, charge transfer, and dissociative electron attachment. Current astrochemical modelling of interstellar clouds, such as IRC+10216, typically underestimate the observed abundances of CH$_4^-$ ~\cite{mil17,cor13,her08}. Therefore, it has been proposed that ion-neutral reactions
\begin{equation}
\text{C}_{2n}\text{H}^- + \text{C}_2\text{H}_2 \rightarrow \text{C}_{2n+2}\text{H}^- + \text{H}_2.
\end{equation}
may be significant chain growth mechanisms that also needs to be included in astrochemical models~\cite{mil17,bas19}. Studies of the extraterrestrial planetary atmosphere of Saturn's moon Titan suggest that both C$_2$H and C$_2$H$^-$ play an important role in the atmospheric chemistry~\cite{dob16,vui09,des17,vri18}. Multiple reaction pathways involving the C$_2$H$^-$ have been proposed, including reactions with C$_2$H$_2$ and HCN.

@@

In this work the vibronic spectrum of the ethynyl radical C$_2$H is examined with High-Resolution Photoelectron Imaging (HR-PEI) together with CFOUR~\cite{dev20}  \emph{ab-initio} vibronic-coupling calculations. C$_2$H is reported to be one of the most abundant molecules in the universe, and is the most thoroughly studied of the C$_{2n}$H species~\cite{wil91,hei99,for10}. Many different experimental techniques have been employed to help understand the complexities of the vibronic spectrum, including electron spin resonance~\cite{coc64,jin85}, laser magnetic resonance~\cite{bro88,pfe96,sch98}, microwave and millimeter wave spectroscopy~\cite{sas81,got83,end89,mul00}, infrared (matrix isolation~\cite{she87,jac87,for95} and Fourier Transform emission~\cite{ver88}) spectroscopy, laser induced fluorescence spectroscopy~\cite{hsu93,hsu95,chi99} and photoelectron spectroscopy of the negative ion~\cite{erv91,tay98,zho07}. C$_2$H has also received extensive theoretical attention to interpret the experimental results, however the large number of possible vibronic levels means this has proven to be a challenge~\cite{per90,per91,per91b,per91c,per92,car00,tar03,tar04,for10,sta21,gul21}. In this work we demonstrate how the construction of a quasidiabatic Hamiltonian allows the strength of vibronic interactions between coupled surfaces near a conical intersection to be estimated, sufficient to accurately simulate electronic and vibronic spectra. Anion HR-PEI spectroscopy maps both of the $^2\Sigma^+$ and $^2\Pi$ surfaces on an equal footing from the anion $\tilde{X}^1\Sigma^+$ state, allowing a direct comparison with \emph{ab-initio} modelling.

\section{Methods}
\subsection{Experimental Procedure}
Details of the HR-PEI spectrometer are given in Refs~\onlinecite{cav07} and~\onlinecite{dev17}. Briefly, C$_2$H$^-$ and C$_2$D$^-$ anions are produced by passing pure ethylene (C$_2$H$_4$ or C$_2$D$_4$) gas at $\sim2$ bar pressure through a pulsed solenoid valve, into the source vacuum chamber. The gas undergoes adiabatic supersonic expansion into a high-voltage discharge which generates negative ions. The anions are extracted, accelerated to 500~eV, and focussed into a novel gating, bunching, and re-referencing unit~\cite{ded01}, before being mass separated over a 2m time-of-flight region. The anion of interest (C$_2$H$^-$ or C$_2$D$^-$) is then isolated by an electrostatic gate. The ion packet is crossed with a tuneable detachment laser beam, generated from a Continuum Sunlite EX optical parametric oscillator pumped by the third harmonic of a Continuum Powerlite 9010 Nd:YAG laser. The laser output is then doubled by passing through a BBO crystal, to produce UV laser pulse energies between 1-5mJ at 10Hz. The wavelength of the laser light is measured using a HighFinesse WS7 UV wavemeter.

\subsection{Computational Details}
Quantum chemical computations were preformed using the CFOUR computational package~\cite{dev20}. The equation-of-motion ionised-potential coupled-cluster singles, doubles, and triples (EOM-IP-CCSDT) level of theory with an atomic natural orbital basis set of quadruple-zeta quality (ANO2) was used on a closed-shell anionic reference to perform geometry optimisations and frequency calculations (including 1h + 2h1p + 3h2p excitations~\cite{mat16}). EOM is a highly accurate and robust framework that provides access to different types of target electronic states, including ionised states~\cite{gul21b,and22,bar07}. The linear diabatic coupling constants were calculated analytically at the EOM-IP-CCSD/ANO2 level.

@@

In C$_2$H the loss of degeneracy creates a Renner-Teller (RT) pair in the excited state, where the usually degenerate $\Pi$ surfaces separate to form two non-degenerate electronic states $\Pi^+ (2A')$ and $\Pi^-(1A'')$. This involves separating a single potential energy surface (V) into two distinct but coupled surfaces (V$^+$) and (V$^-$). Due to the strong coupling along the linear axis between the electronic and vibration angular momenta of the $2A'$ and $1A''$ components of the $^2\Pi$ state, stationary states cannot be explicitly assigned to either of the $\Pi^+(2A')$ or $\Pi^-(1A'')$ electronic surfaces. Instead, they exist as a combination of both states. Due to the close lying nature of the ground {$\tilde{X} ^2\Sigma^+$} and excited {$\tilde{A} ^2\Pi$} electronic states, which are only separated by $\sim3700~$cm$^{-1}$, a pseudo Jahn-Teller effect is also observed. In the case of C$_2$H this is seen as coupling between the ground $\Sigma^+(1A')$ and excited $\Pi^+(2A')$ states, induced by the bending motion of $v_2$. The ground state only couples to one member of the Renner-Teller pair $\Pi^+(2A')$, as the other state $\Pi^-(1A'')$ has incompatible symmetry. This results in the complex vibronic structure observed for the $\tilde{A}^2\Pi$ electronic state, with contributions from three coupled surfaces $\Sigma^+(1A'),$ $\Pi^+(2A'),$ and $\Pi^-(1A'')$.

@@

\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth]{figures/Fig2}
	\caption{Adiabatic potential energy surfaces of $\tilde{X}^2\Sigma^+$, $\tilde{A}^2\Pi^-$, and $\tilde{A}^2\Pi^+$ states, calculated using the parameters calculated by Tarroni and Carter~\cite{tar03}.}
	\label{fig:2}
\end{figure}

The adiabatic potential energy surfaces in Figure~\ref{fig:2} were evaluated from the variational parameters of Tarroni and Carter~\cite{tar03}, illustrating the loss of degeneracy for the doublet $^2\Pi$ surfaces, due to the Renner-Teller and pseudo Jahn-Teller interactions. This technique has been used to calculate a multitude of admixed vibronic levels. More than 100 possible C$_2$H vibronic states have been calculated, up to 10,000~cm$^{-1}$ above the $\tilde{X} ^2\Sigma^+$ origin. However, assigning experimental spectra has remained a challenge, partly due to the large number of calculated levels providing for multiple possible assignments for each experimentally observed transition.

\subsection{Coupling Calculations}
In order to guide spectral assignment of C$_2$H and future astronomical searches for other C$_n$H radicals, transition intensities are also needed. However, due to the strong vibronic coupling interactions, transition intensities can not be obtained via simple Franck-Condon calculations based on quantum chemistry methods. To account for these interactions, the photoelectron spectrum of the C$_2$H$^-$ anion is simulated using a quasidiabatic Hamiltonian of the type introduced by Koeppel, Domcke and Cederbaum (KDC).~\cite{kou84,dom81} In this approach, the Hamiltonian is represented in a basis of quasidiabatic (slowly varying) electronic  states for which the kinetic energy operator can be assumed diagonal.~\cite{pac93} For C$_2$H, the KDC Hamiltonian comprises three states - the two components of the $^2\Pi$ state and the ground $^2\Sigma^+$ state - and is then projected onto a vibrational basis, usually chosen as a direct product of harmonic oscillators. Diagonalization of the corresponding matrix yields the molecular states (which are given in terms of a Born-Huang expansion), and the squared projections of the corresponding eigenvectors onto the ground state of the anion yield the relative intensities. The latter is true only if the photodetachment cross sections for the $^2\Sigma^+$ and $^2\Pi$ states are assumed equal, but different cross sections of the two states can be incorporated by scaling the intensities of states according to their (vibronic) symmetry.~\cite{sta11}

@@

When the potential in Eq.~(\ref{eq:calc1}) is diagonalized, the adiabatic states that are used for its parametrization are precisely recovered through terms second order in displacement. The parametrization of the potential in this work was completed using the CFOUR computational package~\cite{dev20} and the quasidiabatic ansatz of Ichino~\emph{et al.}~\cite{ich09} Briefly,the EOM procedure is used to operationally define quasidiabatic states (those that relax according to a well-behaved reference state wavefunction, which is that of the anion level) and the coupling constants are then evaluated as the first derivative of the off-diagonal elements of the electronic Hamiltonian in the basis defined by this representation. Vertical excitation energies ($\Delta_0$), and linear diabatic force and coupling constants between the electronic states of interest were calculated analytically using the highly accurate EOM-IP-CCSDT/ANO2 procedure. The interstate coupling constants ($\lambda$) were calculated analytically at the EOM-IP-CCSD/ANO2 level. All calculated values used to parametrize the KDC potential in Eq.~(\ref{eq:calc1}) are presented in Table~\ref{tab:1}.

\begin{table}
	\caption{Parameters for the quasidiabatic Hamiltonian Eq.~(\ref{eq:calc1}) determined using CFOUR and the quasidiabatic ansatz. Units are in cm$^{-1}$.}
	\label{tab:1}

	\begin{tabular}{c | r r r r | r r r r} 
		\hline
		& \multicolumn{4}{c}{C2H} & \multicolumn{4}{c}{C2D}\\
		\hline
		& ~~Anion~~ & ~~$\tilde{X}^2\Sigma^+$~~ & ~~$\tilde{A}^2\Pi (a')$~~ & ~~$\tilde{A}^2\Pi (a'')$~~ & ~~Anion~~ & ~~$\tilde{X}^2\Sigma^+$~~ & ~~$\tilde{A}^2\Pi (a')$~~ & ~~$\tilde{A}^2\Pi (a'')$~  \\
		\hline
		$\Delta_0$ & 0& 22431.21&  26543.67& 26543.67& 0 & 22431.21& 26543.67& 26543.67 \\
		F$_1$ &  0& -211.28& 286.90& 286.90& 0& -596.04& 641.6& 641.6\\
		F$_3$ &  0& -1515.21&  1257.07& 1257.07& 0& -1401.53& 1115.17& 1115.17\\
		F$_{2a}$& 0 & 0& 0& 0& 0& 0& 0&0 \\
		F$_{2b}$&  0& 0& 0& 0& 0& 0& 0&0 \\
		F$_{11}$& 3424.32 &  3487.03& 3482.16& 3482.16& 2611.42&  2636.57& 2681.25& 2681.25\\
		F$_{13}$&  0 & -28.32& 31.46& 31.46& 0& -48.38& 55.40& 55.40\\
		F$_{33}$& 1894.71 & 1830.83& 2042.21& 2042.21& 1792.49&  1747.62& 1914.39& 1914.39\\
		f$_{22}$& 607.09&  87.57& -& -& 481.89&  120.09& -& -\\
		F$_{22}$& 607.09 & 770.78& -& -& 481.89& 620.70& -& -\\
		f$_{2a2a}$& - &- & 1883.48& 658.96& -& -& 1503.88& 473.55\\
		f$_{2b2b}$& - &- & 658.96& 1883.48& -& -& 473.55& 1503.88\\
		F$_{2a2a}$& -& -& 1200.27& 658.96& -& -& 1003.27& 473.55\\
		F$_{2b2b}$& -& -& 658.96& 1200.27& -& -& 473.55& 1003.27\\
		&&&&&&&& \\
		$\lambda$& & & -1185.26& & & & -1014.58& \\
		$\eta$ & & & -270.66& & & & -264.86& \\

@@

\subsection{Simulated Spectra}
The above method was employed to simulate the photoelectron spectrum of C$_2$H$^-$, with the calculated transitions shown in Figure~\ref{fig:3}(a) alongside the experimental data at 355~nm. Photoelectron spectra are simulated using the xsim package of CFOUR~\cite{dev20} and the model Hamiltonian generated above. The xsim module projects the Hamiltonian (Eq.~\ref{eq:calc1}) onto a vibrational basis, which is then diagonalised using Lanczos algorithm to calculate transition energies and intensities that map to the measured photoelectron spectrum. In this simulation the transition moments for the two ionisation processes are assumed to be equal.   
On the $\tilde{X} ^2\Sigma^+$ surface, below 3,600~cm$^{-1}$, there is excellent agreement in both the transition positions and intensities between the simulated and experimental spectra. This includes the HT coupled (2$^{n+1}$) transitions with $\Pi$ symmetry, which would normally be missing from a simulation using the standard \emph{ab-initio} approaches.

@@

\begin{figure}[th!]
	\includegraphics[width=0.6\textwidth]{figures/Fig3}
	\caption{(a) Experimental photoelectron spectrum of C$_2$H$^-$ at 355~nm, compared with the simulated spectrum, calculated using the quasidiabatic ansatz. Transitions with $\Sigma$ symmetry are shown in blue, and those with $\Pi$ symmetry are shown in orange. (b) Experimental photoelectron spectrum of C$_2$D$^-$ at 355~nm, compared to the simulated spectrum. The labelling of peaks near the $\tilde{A}$ origin correspond to transitions identified in Tables~S1 and~S2.}
	\label{fig:3}
\end{figure}

@@

The simulated spectrum reproduces the large change observed for the deuterated species. The calculated intensity pattern of the vibronic levels around the $\tilde{A} ^2\Pi$ origin also mirrors the experimental data, with a single intense transition observed near peak $e$, and 3 prominent transitions predicted around the experimental peaks $a-d$. The calculated splitting between the levels is also similar.

@@
